---
title: "ISLR Chapter 10"
author: "Alex Bayers"
date: "11/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exercise 2
```{r}
data = c(0, .3, .4, .7, 0.3, 0, 0.5, 0.8, 0.4, 0.5, 0, 0.45, 0.7, 0.8, 0.45, 0)
data <- as.dist(matrix(data, nrow = 4, ncol = 4), diag = TRUE)
complete.clust <- hclust(data)
single.clust <- hclust(data, method = "single")
complete.trees <- cutree(complete.clust, k = 2)
plot(complete.clust, main = "Complete Linkage", xlab = "")
abline(a=.5, b = 0, col = "red")
plot(single.clust, main = "Single Linkage", xlab = "")
abline(a=.42, b = 0, col = "red")
```
Complete linkage results in clusters of 1 & 2, and 3 & 4.  Single linkage results in 1, 2, 3, and a separate cluster at 4.

###Exercise 3
```{r}
library(ISLR)
set.seed(4)
x1 <- c(1, 1, 0, 5, 6, 4)
x2 <- c(4, 3, 4, 1, 2, 0)
ylabel <- sample(c(1, 2), 6, TRUE)
my.frame <- data.frame(x1, x2, ylabel)
oldylabel  <- rep(0, 6)
plot(x = x1, y = x2, col = ylabel, pch = 16)

while (any(oldylabel != my.frame$ylabel)){
  y1.mean <- colMeans(subset(my.frame, ylabel == 1))
  y2.mean <- colMeans(subset(my.frame, ylabel == 2))
  
  y1.demeaned <- sweep(my.frame, 2, y1.mean)
  y2.demeaned <- sweep(my.frame, 2, y2.mean)
  
  x1.distance <- (y1.demeaned$x1^2 + y1.demeaned$x2^2)^.5
  x2.distance <- (y2.demeaned$x1^2 + y2.demeaned$x2^2)^.5
  oldylabel <- my.frame$ylabel
  my.frame$ylabel <- ifelse(x1.distance < x2.distance, 1, 2)
  newtest = 1
}
plot(x = x1, y = x2, col = my.frame$ylabel, pch = 16)
print(my.frame)
```

###Exercise 4
#####A
The fusion will occur higher up in the complete linkage rather than the single linkage, because they will be fused on the maximal distance rather than the minimum distance.

#####B
They will fuse at the same height, because there are no other linkages when they're fusing.

#Exercise 5.
Left Hand Chart - We would group yellow, purple, black, and light red together as one group.
Center Chart - We would group yellow, blue, light red, and purple together.
Right Chart - We would group yellow, blue, light red, and purple together.

###Exercise 7
```{r}
library(ISLR)
scaled.arrests <- scale(USArrests)
arrests.dist <- dist(scaled.arrests, diag = TRUE)
arrests.cor <- cor(t(scaled.arrests))
#summary((arrests.dist ^ 2) / (1 - arrests.cor))
```

###Exercise 8
```{r}
scaled.arrests <- scale(USArrests)
#A
pc.arrests <- prcomp(scaled.arrests)
pc.var <- pc.arrests$sdev^2
pve <- pc.var/sum(pc.var)
pve

#B?
sum(scaled.arrests^2)
```

